Best practices for forecasting changes in biodiversity: an example using breeding birds

### 2017-09-05 manuscript/draft.Rmd
fix timescale discussion based on our meeting

We have only evaluated annual forecasts up to a decade into the future, but
forecasts are often made with a lead time of 50 years or more. These long-term
forecasts are difficult to evaluate given the small number of century-scale
datasets, but are important for understanding changes in biodiversity at some of
the lead times relevant for conservation and management. Two studies have
assessed models of species richness at longer lead times [@algar2009;
@distler2015], but the results were not compared to baseline or time-series
models (in part due to data limitations) making them difficult to compare to our
results directly. Given the potential changes with lead time of the models and
processes are most relevant to forecasting, it is important that more forecasts
with longer lead times are evaluated and that the inferences drawn from
short-term and long-term forecasting studies be compared. By evaluating
shorter-term forecasts, researchers can avoid waiting so many decades to
evaluate their methods [@petchey2015; @tredennick2016; @dietze2016]. However,
since drivers of species richness differ at different temporal scales
[@rosenzweig1995; @white2004; @white2007], short-term evaluations should be
viewed as informing, not replacing, long-term evaluations [@tredennick2016]. In
particular, we might expect environmental and ecological information to become
more important at longer time scales, and thus for the performance of simple
baseline forecasts to degrade faster than forecasts from SDMs and other similar
models [@blonder2017].  On the other hand, there is evidence that species
respond differently to spatial versus temporal variation in the environment
[@oedekoven2017], implying that most SDMs are not using environmental
information well enough to make accurate forecasts.

>>>>

We have only evaluated annual forecasts up to a decade into the future, but
forecasts are often made with a lead time of 50 years or more. These long-term
forecasts are difficult to evaluate given the small number of century-scale
datasets, but are important for understanding changes in biodiversity at some of
the lead times relevant for conservation and management. Two studies have
assessed models of species richness at longer lead times [@algar2009;
@distler2015], but the results were not compared to baseline or time-series
models (in part due to data limitations) making them difficult to compare to our
results directly. Studies on shorter time scales, such as ours, provide one way
to evaluate our forecasting methods without having to wait several decades to
observe the effects of environmental change on biodiversity [@petchey2015;
@tredennick2016; @dietze2016], but cannot fully replace longer-term evaluations
[@tredennick2016]. In general, drivers of species richness can differ at
different temporal scales [@rosenzweig1995; @white2004; @white2007], so
different methods may perform better for different lead times. In particular, we
might expect environmental and ecological information to become more important
at longer time scales, and thus for the performance of simple baseline forecasts
to degrade faster than forecasts from SDMs and other similar models
[@blonder2017].  We did observe a small trend in this direction: deviance for
the auto-ARIMA models and for the average baseline grew faster than for two of
the environmental models (the JSDM and the macroecological model), although this
growth was not statistically significant for the average baseline.
